Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys

### 2019-01-25 doc/article.tex
going on

We consider peculiar velocity surveys for a range of redshift depths $z_{max}$ for survey durations of $t=2$ and 10~years.
The other survey parameters
$\Omega=3\pi$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
The $k$-limits are  taken to be $k_{min} = (2 r_{max})^{-1}$  and 
$k_{max} = 0.1$~$h$\,Mpc$^{-1}$.
The input bias of SN~Ia host galaxies is set as $b=1.2$.

@@

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/var.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{The projected uncertainty in $\gamma$ achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed) , one that uses
both overdensities (RSD) and peculiar velocities independently (dotted), and one  that uses
both RSD, peculiar velocities, and their cross-correlation (solid).
\label{var:fig}}
\end{figure}

Survey performance is examined in more detail by considering how $\sigma_\gamma$ changes with respect to the 
survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
within a given survey.  
\begin{description}
\item[Solid Angle $\Omega$]
The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole could significantly increase
the sky coverage with a corresponding decrease in the variance of $\gamma$.
\item[Differential Redshift Bin $z$] Certain differential redshift bins  constrain $\gamma$ more strongly than others.  
Figure~\ref{dvar:fig} shows $\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$.
In both cases the shape of the curve is the same, with the amplitude monotonically decreasing from $z=0.01$ out to $z=0.2$.  If we had
to sculpt the distribution (say due to limited follow-up resources), the preference would be to sculpt out the highest redshift bins
resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the SN-discovery distribution and the sample is unambiguously
parameterized by $z_{max}$.  Note that this result is based on constraining a constant redshift-independent $\gamma$.
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{src/dvardz.png}
%\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{$\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$. 
\label{dvar:fig}}
\end{figure}
\item[Redshift Depth $z_{max}$]
Increasing the survey redshift depth increases the $\gamma$ precision though that precision asymptotes
$z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
 \item[Survey duration $t$] There is a significant decrease on $\sigma_\gamma$ when going from a 2- to ten-year survey, as shown
 in  Figure~\ref{var:fig}.   The left panel of Figure~\ref{dvar2:fig} shows $\partial \sigma_\gamma / \partial \ln{t}$ as a function of
 $z_{max}$ for two- and ten-year surveys.  The non-zero amplitudes show that shot-noise affects both two- and ten-year surveys, but
 much less so for the latter more so at low redshifts.  After 10-years, the effect of sample-variance important for shallower surveys.
 \begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
\includegraphics[width=0.47\textwidth]{src/dvardsigM.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Left: $\partial \sigma_\gamma / \partial \ln{t}$  and Right: $\partial \sigma_\gamma / \partial \sigma_M$ as a function of
 $z_{max}$ for two- and ten-year surveys.
\label{dvar2:fig}}
\end{figure}
\item[Intrinsic Magnitude Dispersion $\sigma_M$]
The relationship between intrinsic magnitude dispersion and precision in $\gamma$ can be seen through the 
previous result, as $\partial \sigma_\gamma/\partial \sigma_M = \partial \sigma_\gamma / \partial \ln{t}$.
Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
Thus, a decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
%limited so the $\sigma^2_M/n$-term
%in the Fisher matrix is important.
Survey follow-up strategy,
which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or equivalently in survey duration. 
 \end{description}

Shot-noise limit touched upon several times in the preceding discussion.  Go back to Figure.

There are a number of interesting features to note:
\begin{itemize}
\item Ten-year survey has lower  $\sigma_\gamma$  relative to  two-year surveys with factors that are redshift-dependent.  For $z_{max}=0.3$
the uncertainties are reduced by a factor $\sim 0.5$ whereas at  $z_{max}=0.7$ the factor is $\sim 0.7$
\end{itemize}

 Shot noise
is important  Figure~\ref{power:fig}.
 (the total rest-frame volumetric SN~Ia rate being $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.)

@@

%
%\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
%\subsection{Fisher Matrix of Velocity-Galaxy Surveys}
%
%
%
%\citet{2017ApJ...847..128H, 2017MNRAS.464.2517H} project the precision in $f\sigma_8$, summarized in   Table~\ref{tab:howlett},
% obtained by  complementary galaxy and peculiar velocity surveys in a common volume.
%Their analysis is performed in Fourier space, 
%where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero is given by
%\begin{align}
%F_{ij} & = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
%\frac{\partial C}{\partial \lambda_j} \right]\\
%& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
%\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
%\label{fisher:eqn}
%\end{align}
% and covariance
%\begin{equation}
%C(k,\mu)  =
%  \begin{bmatrix}
%   P_{gg}(k,\mu) + \frac{1}{n_g} &
%   P_{vg}(k,\mu)  \\
%   P_{vg}(k,\mu)  &
%  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
%   \end{bmatrix}.
%\label{cov:eq}
%\end{equation}
%
%
%The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
%\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the galaxy bias.
%The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the per-probe velocity dispersion
%is $\sigma$. 
%The integrand in the Fisher-element calculation
%\begin{equation}
%I = \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial (f\sigma_8)} C^{-1}
%\frac{\partial C}{\partial  (f\sigma_8)} \right]
%\end{equation}
%represents the contribution to  $F_{f\sigma_{8}\,f\sigma_{8}}$ from a
%differential Fourier
%volume element.  
%
%%
%%\begin{figure}
%%\centering
%%\includegraphics[width=0.5\textwidth]{src/new.png}
%%%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
%%\caption{ $I/I_{max}$ for $\mu=1, 0.5$  where $I=\text{Tr}\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$ represents the contribution of a single mode to
%%the Fisher element $F_{f\sigma_8 f\sigma_8}$ and $I_{max}$ is the sample variance limit.  The dashed black curve represents galaxy and velocity surveys in
%%independent volumes.  Two- and ten-year SN number densities are given by dotted and solid color lines respectively.  Different colors represent varying levels
%%of intrinsic velocity dispersion.  The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
%%redshift depths $z_{max}$.
%%\label{shot:fig}}
%%\end{figure}
%
%
%\subsection{Precision in $f\sigma_8$ and Survey Design Parameters}
%First, we demonstrate that the realistic surveys being considered have significant shot noise contributions in their error budgets. 
%
%The $I/I_{max}$ of these  scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
%Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ and so are sample-noise limited.
%Individual $k>0.01$ modes have  $I/I_{max}<1$ meaning  that shot noise
%is significant.  Recall that collectively, $k>0.01$ modes are the important contributors to the peculiar-velocity signal.
Basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$.
%The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).
%\subsection{Survey Parameters that Determine Scientific Performance}

@@

\begin{itemize}
\item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively, we consider
a pure SN~Ia sample obtained through
spectroscopic classification; the requisite follow-up resources needed for complete $z<0.3$ coverage is conceivable. 
The feasibility of using photometric classification is an important subject of research
that can have implications for survey planning, particularly in extending the redshift range.
\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  In most cases this redshift will be
available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
\item Role of LSST:
LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
The nominal cadence  
 produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
The wide-field of LSST is
 well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
$z<0.3$ SNe~Ia.  LSST is suited for maximizing $\Omega$ while other (cheaper) resources that supplement LSST photometry can improve $\sigma_M$.
The ultimate WFD observing strategy determines how well $\sigma_8$ can be determined with LSST data alone, and the amount of external resources needed  (if any) to
reach a targeted $\sigma_M$.
\item Follow-up resources to determine per-SN distances:
Supplemental non-LSST follow-up data in the form of improved temporal light-curve sampling (accurate rise and decline times), 
expanded (UV, NIR) wavelength coverage, and spectral features
can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
 \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.

@@

>>>>

We consider peculiar velocity surveys for a range of redshift depths $z_{max}$ for survey durations of $t=2$ and 10~years.
The other survey parameters
$\Omega=3\pi$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
The $k$-limits are  taken to be $k_{min} = (2 r_{max})^{-1}$  and 
$k_{max} = 0.1$~$h$\,Mpc$^{-1}$.
The sample-selection efficiency $\epsilon$ is redshift-independent, i.e.\ the native redshift distribution is not sculpted.
The input bias of SN~Ia host galaxies is set as $b=1.2$.

@@

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/var.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{The projected uncertainty in $\gamma$ achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed), one that uses
both overdensities (RSD) and peculiar velocities independently (dotted), and one  that uses
both RSD, peculiar velocities, and their cross-correlation (solid).
\label{var:fig}}
\end{figure}

Survey performance is examined in more detail by considering how $\sigma_\gamma$ (using RSD, peculiar velocities, and their cross-correlations) 
changes with respect to the 
survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
within a given survey.  
\begin{description}
\item[Solid Angle $\Omega$]
The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \sqrt{\Omega}$.
\item[Differential Redshift Bin $z$]  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
This is demonstrated to be the case at the end of both 2- and 10-year surveys and $z_{max}=0.2$.
Figure~\ref{dvar:fig} shows
 $|\partial \sigma_\gamma / \partial z|$, which
for both surveys monotonically decreases from $z=0.01$ out to $z=0.2$.  If we had
to sculpt the distribution (say due to limited follow-up resources), the preference would be to cut out the highest redshift bins
resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{max}$.
\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/dvardz.png}
%\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{$|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
\label{dvar:fig}}
\end{figure}
\item[Redshift Depth $z_{max}$]
Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
$z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
 \item[Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$] Increasing the survey duration to increase the number density improves constraints.  The benefit is stronger for
shallow compared to deep surveys, and the strength of the  benefit drops significantly over the course of 10 years.
These trends are seen in the left panel of Figure~\ref{dvar2:fig}, which plots $|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
 $z_{max}$ for two- and ten-year surveys, and
reflect those $k$-modes that do not benefit more data because they are sample-variance limited, either because they are at low $k$ for shallow surveys
or have already been saturated over the course of the survey.
 \begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
\includegraphics[width=0.47\textwidth]{src/dvardsigM.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Left: $|\partial \sigma_\gamma / \partial \ln{t}|$  and Right: $\partial \sigma_\gamma / \partial \sigma_M$ as a function of
 $z_{max}$ for two- and ten-year surveys.
\label{dvar2:fig}}
\end{figure}
Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
has similar relationships with $\sigma_\gamma$.
This is evident in the plot of $\partial \sigma_\gamma/\partial \sigma_M$ shown in the right panel of Figure~\ref{dvar2:fig}.
The improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or equivalently in survey duration. 
 \end{description}

From the above calculations, we find empirically that shot noise is important since higher number densities and lower intrinsic magnitude dispersion
significantly reduce $\gamma$ uncertainty, but that sample variance is also relevant  by the end of a ten-year survey.
Survey design is aided by understanding the circumstances in which  the surveys become sample variance limited.
First note that the  rest-frame volumetric SN~Ia rate (full selection efficiency) is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$;
after a few years the RSD measurement is sample variance limited.  The  contributions of $P_{vv}$ and
$\sigma^2/n$ at $z=0.1$ for $\sigma_M=0.08, 0.15$~mag with number densities from 2- and 10-year surveys
sare shown in  Figure~\ref{power:fig}.  For all of these survey scenarios, the transition between sample variance and shot-noise equality 
occurs at a $k$ squarely within the $k$-range that carries the most signal.

@@

%
%\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
%\subsection{Fisher Matrix of Velocity-Galaxy Surveys}
%
%
%
%\citet{2017ApJ...847..128H, 2017MNRAS.464.2517H} project the precision in $f\sigma_8$, summarized in   Table~\ref{tab:howlett},
% obtained by  complementary galaxy and peculiar velocity surveys in a common volume.
%Their analysis is performed in Fourier space, 
%where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero is given by
%\begin{align}
%F_{ij} & = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
%\frac{\partial C}{\partial \lambda_j} \right]\\
%& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
%\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
%\label{fisher:eqn}
%\end{align}
% and covariance
%\begin{equation}
%C(k,\mu)  =
%  \begin{bmatrix}
%   P_{gg}(k,\mu) + \frac{1}{n_g} &
%   P_{vg}(k,\mu)  \\
%   P_{vg}(k,\mu)  &
%  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
%   \end{bmatrix}.
%\label{cov:eq}
%\end{equation}
%
%
%The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
%\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the galaxy bias.
%The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the per-probe velocity dispersion
%is $\sigma$. 
%The integrand in the Fisher-element calculation
%\begin{equation}
%I = \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial (f\sigma_8)} C^{-1}
%\frac{\partial C}{\partial  (f\sigma_8)} \right]
%\end{equation}
%represents the contribution to  $F_{f\sigma_{8}\,f\sigma_{8}}$ from a
%differential Fourier
%volume element.  
%
%%
%%\begin{figure}
%%\centering
%%\includegraphics[width=0.5\textwidth]{src/new.png}
%%%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
%%\caption{ $I/I_{max}$ for $\mu=1, 0.5$  where $I=\text{Tr}\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$ represents the contribution of a single mode to
%%the Fisher element $F_{f\sigma_8 f\sigma_8}$ and $I_{max}$ is the sample variance limit.  The dashed black curve represents galaxy and velocity surveys in
%%independent volumes.  Two- and ten-year SN number densities are given by dotted and solid color lines respectively.  Different colors represent varying levels
%%of intrinsic velocity dispersion.  The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
%%redshift depths $z_{max}$.
%%\label{shot:fig}}
%%\end{figure}
%
%
%\subsection{Precision in $f\sigma_8$ and Survey Design Parameters}
%First, we demonstrate that the realistic surveys being considered have significant shot noise contributions in their error budgets. 
%
%The $I/I_{max}$ of these  scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
%Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ and so are sample-noise limited.
%Individual $k>0.01$ modes have  $I/I_{max}<1$ meaning  that shot noise
%is significant.  Recall that collectively, $k>0.01$ modes are the important contributors to the peculiar-velocity signal.
%Basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$.
%The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).
%\subsection{Survey Parameters that Determine Scientific Performance}

@@

\begin{itemize}
\item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively, we consider
a pure SN~Ia sample obtained through
spectroscopic classification; the requisite follow-up resources needed for complete $z<0.3$ coverage is conceivable. 
The feasibility of using photometric classification is an important subject of research
that can have implications for survey planning, particularly in extending the redshift range.
\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  In most cases this redshift will be
available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
\item Role of LSST:
LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
The nominal cadence  
 produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
The wide-field of LSST is
 well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
$z<0.3$ SNe~Ia.  LSST is suited for maximizing $\Omega$ while other (cheaper) resources that supplement LSST photometry can improve $\sigma_M$.
The ultimate WFD observing strategy determines how well $\sigma_8$ can be determined with LSST data alone, and the amount of external resources needed  (if any) to
reach a targeted $\sigma_M$.

The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole could significantly increase
the sky coverage with a corresponding decrease in the variance of $\gamma$.

\item Follow-up resources to determine per-SN distances:
Supplemental non-LSST follow-up data in the form of improved temporal light-curve sampling (accurate rise and decline times), 
expanded (UV, NIR) wavelength coverage, and spectral features
can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
 \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.

@@

Effect of 10-years.
